---
title: "FCAnalysisWPLI"
output: powerpoint_presentation
date: "2024-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r libraries}
library(tidyverse)
library(lattice)
library(latticeExtra)
library(gridExtra) # adjust plots in a grid
library(lme4) # linear mixed models
library(kableExtra)
library(nlme)
library(performance)
library(emmeans) # post hoc multiple comparisons of linear mixed models
library(see)
library(patchwork)
library(ggplot2)
library(kableExtra)
library(rstatix)
```

```{r}
setwd("W:/Projects/2019-04 M1M1PAS Project/analysis/M1M1PASgit")

FCdata <- read.csv('W:/Projects/2019-04 M1M1PAS Project/analysis/source_Paolo/fc_source/sourceConnectivityBeamformeroldLF.csv', header=TRUE, sep=',', dec='.')
FCdata$Subject <- factor(FCdata$subject)
FCdata$Intervention <- factor(FCdata$task,levels=c("random","negneg","negpos","posneg"))
FCdata$Time <- factor(FCdata$run,levels=c("2","1","3","4","5"))
FCdata$wPLI <- as.numeric(FCdata$wpli)
FCdata$AEC <- as.numeric(FCdata$aec)
```

```{r}

FCdata$Response <- FCdata$wPLI
#FCdata$Response <- FCdata$AEC

```

## Check distribution of the dependant variable

The wPLI is not normally distributed.

```{r, fig.dim = c(15,10)}

xy.time <- xyplot(Response ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "FC", lwd = 2)

xy.interv <- xyplot(Response ~ Intervention, data = FCdata, type = c("a", "p"), pch = 20,
                 group = Subject, xlab = "Intervention", ylab = "FC", lwd = 2)
                    
qq.data <- qqmath(~ Response, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FCdata, aes(x = Response)) +
  geom_histogram(binwidth = 0.1) +
  xlab("FC") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,qq.data,hist.data), ncol = 2, nrow = 2)
```

## Check distribution of the normalised variable

We transform it by taking an atanh and then taking it to the power of 0.24. If AEC, then just the power pf 0.30

```{r}
FCdata$ResponseNorm <- atanh(FCdata$Response)^0.24
#FCdata$ResponseNorm <- (FCdata$Response)^0.30
```

```{r, fig.dim = c(15,10)}

xy.time <- xyplot(ResponseNorm ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "Normalised FC", lwd = 2)

xy.interv <- xyplot(ResponseNorm ~ Intervention, data = FCdata, type = c("a", "p"), pch = 
                      20, group = Subject, xlab = "Intervention", ylab = "Normalised FC", 
                    lwd = 2)
                    
qq.data <- qqmath(~ ResponseNorm, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FCdata, aes(x = ResponseNorm)) +
  geom_histogram(binwidth = 0.1) +
  xlab("Normalised FC") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,qq.data,hist.data), ncol = 2, nrow = 2)

```

## LME. Model selection

A model with a session as a random effect (1 \| Subject/Intervention) has lower AIC and BIC values, so we will take it for further analysis.

```{r}
model1 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

model2 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

anova(model1,model2)
```

## LME. Significance test

2-way interaction of Intervention:Time is significant. The main effect of Time is also significant.

```{r}
model <- lmerTest::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
car::Anova(model)
```

## LME. Model quality checks

Model residuals are approximately okay but collinearity is very high.

```{r, fig.dim = c(15,30)}
#check_model(model)
```

## LME. Model quality checks

```{r, fig.dim = c(8,8)}

# # Extract residuals and fitted values
# residuals <- resid(model)
# fitted <- fitted(model)
# 
# # Plot residuals vs. fitted values
# ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = Residuals)) +
#   geom_point() +
#   geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
#   labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals")

```

## LME. Model quality checks

```{r, fig.dim = c(8,8)}
# ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = sqrt(abs(Residuals)))) +
#   geom_point() +
#   geom_smooth(method = "loess", color = "red") +
#   labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Square Root of |Residuals|")
```

## LME. Model quality checks

If we fit a model without interactions, with just main effects, there is no collinearity. So we will ignore this issue.

```{r}
# model_reduced <- lmerTest::lmer(ResponseNorm ~ Intervention + Time + (1|Subject/Intervention),
#                      data = FCdata, REML = T,
#                      control = lmerControl(optCtrl = list(maxfun = 1e6)))
# check_collinearity(model_reduced)
```

## Post-hoc plots of significant effects

We will plot not raw but normalised data, the way it was used in the model. Since it's approximately normal, the y-axis is mean and SE of the data.

```{r, fig.dim = c(10,8)}

sub_sum_FCdata <- FCdata %>%   
  group_by(Intervention, Time, Subject) %>%  
  summarize(FC_normed_mean = mean(ResponseNorm)) %>%
  ungroup()

sum_FCdata <- sub_sum_FCdata %>%   
  group_by(Intervention, Time) %>%  
  summarize(ResponseNorm_mean = mean(FC_normed_mean),
            ResponseNorm_se = sd(FC_normed_mean)/sqrt(length(FC_normed_mean))) %>%
  ungroup()

sum_FCdata %>%
  ggplot(aes(x = Time,  
             y = ResponseNorm_mean,
             group = Intervention,  
             color = Intervention)) +    
  geom_point(size = 3) + 
  geom_line(size=1) +
  ylab("Mean normalised FC") +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se, ymin=ResponseNorm_mean-ResponseNorm_se), width=.1, size = 1) +
  theme(legend.position="top")
```

## Model prediction

Values predicted by the model and confidence limits.

```{r, echo=FALSE}
emm.model <- emmeans(model, ~ Intervention * Time,  data = FCdata)

emmip_output <- emmip(emm.model, ~ Intervention*Time ,type = "response", data = FCdata, CIs = TRUE, plotit=FALSE)

ggplot(data = emmip_output, aes(x=Time, y=yvar,group = Intervention,  
             color = Intervention)) +
  geom_point(size = 3) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=.1, size = 1) +
  ylab("Predicted response") +
  theme(legend.position="top")

```

## Post-hoc pairwise comparisons

All 70 tests were performed with FDR correction. Here only the significant ones are shown.

```{r}
emm_options(lmer.df = 'asymptotic') # also possible: 'satterthwaite', 'kenward-roger', 'asymptotic'

emm.model <- emmeans(model, ~ Intervention * Time, data = FCdata)

emm.model.contrast = contrast(emm.model, "trt.vs.ctrl1", simple = "each",
                                   combine = TRUE, adjust = "fdr")

contrast.summary <- summary(emm.model.contrast,by = NULL)
contrast_pretty <- kable(get_anova_table(contrast.summary), digits = 3)
write.csv(contrast.summary, file = "contrasts_FC_interaction.csv")
contrast_pretty
save_kable(contrast_pretty,"contrasts_FC_interaction.txt")
```

## Post-hoc plots of significant effects

Results may be misleading due to involvement in interactions. Means and SE of normalised FC.

```{r, fig.dim = c(10,8)}
sub_sum_noInt_FCdata <- FCdata %>%   
  group_by(Subject, Time) %>%  
  summarize(ResponseNorm_subject_mean = mean(ResponseNorm)) %>%
  ungroup()

sum_noInt_FCdata <- sub_sum_noInt_FCdata %>%   
  group_by(Time) %>%  
  summarize(ResponseNorm_mean = mean(ResponseNorm_subject_mean),
            ResponseNorm_se = sd(ResponseNorm_subject_mean)/sqrt(length(ResponseNorm_subject_mean))) %>%
  ungroup()

sum_noInt_FCdata %>%
  ggplot(aes(x = Time,  
             y = ResponseNorm_mean)) +    
  geom_point(size = 3) + 
  geom_line(group = 1, size = 1) +
  ylab("Mean normalised FC") +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se, ymin=ResponseNorm_mean-ResponseNorm_se), width=.1, size = 1) +
  theme(legend.position="top")
  
```

## Model prediction

Results may be misleading due to involvement in interactions. Values predicted by the model and confidence limits.

```{r, fig.dim = c(10,8)}

emmip_output <- emmip(emm.model, ~ Time ,type = "response", data = FCdata, CIs = TRUE, plotit=FALSE)

ggplot(data = emmip_output, aes(x=Time, y=yvar)) +
  geom_point(size = 3) + 
  geom_line(group = 1, size = 1) +
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=.1, size = 1) +
  ylab("Predicted response") +
  theme(legend.position="top")

```
