---
title: "FC analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r libraries}
library(tidyverse)
library(lattice) #plotting
library(gridExtra) #plotting
library(lme4) #LME
library(performance) # model evaluation
library(emmeans) #pairwise comparisons
library(kableExtra) #pretty tables
library(rstatix) #pretty tables
```

```{r}
setwd("W:/Projects/2019-04 M1M1PAS Project/analysis/M1M1PASgit")

FCdata <- read.csv('sourceFC.csv', header=TRUE, sep=',', dec='.')
FCdata$Subject <- factor(FCdata$subject)
FCdata$Intervention <- factor(FCdata$task,levels=c("random","negneg","negpos","posneg"))
FCdata$Time <- factor(FCdata$run,levels=c("2","1","3","4","5")) # !!!Baseline 2 is ref, it comes first in plots
FCdata$wPLI <- as.numeric(FCdata$wpli)
FCdata$AEC <- as.numeric(FCdata$aec)
FCdata$Response <- FCdata$wPLI
```

## Check distribution of the dependant variable

```{r, fig.dim = c(15,10)}

xy.time <- xyplot(Response ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "FC", lwd = 2)

xy.interv <- xyplot(Response ~ Intervention, data = FCdata, type = c("a", "p"), pch = 20,
                 group = Subject, xlab = "Intervention", ylab = "FC", lwd = 2)
                    
qq.data <- qqmath(~ Response, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FCdata, aes(x = Response)) +
  geom_histogram(binwidth = 0.1) +
  xlab("FC") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,qq.data,hist.data), ncol = 2, nrow = 2)
```

## Check distribution of the normalised variable

```{r}
FCdata$ResponseNorm <- atanh(FCdata$Response)^0.24

xy.time <- xyplot(ResponseNorm ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "Normalised FC", lwd = 2)

xy.interv <- xyplot(ResponseNorm ~ Intervention, data = FCdata, type = c("a", "p"), pch = 
                      20, group = Subject, xlab = "Intervention", ylab = "Normalised FC", 
                    lwd = 2)
                    
qq.data <- qqmath(~ ResponseNorm, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FCdata, aes(x = ResponseNorm)) +
  geom_histogram(binwidth = 0.1) +
  xlab("Normalised FC") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,qq.data,hist.data), ncol = 2, nrow = 2)

```

## LME. Model selection

```{r}
model1 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

model2 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

anova(model1,model2)
```

## LME. Significance test

```{r}
model <- lmerTest::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
car::Anova(model)
```

## LME. Model quality checks

```{r, fig.dim = c(15,30)}
#check_model(model)
```

## LME. Model quality checks (Linearity)

```{r, fig.dim = c(8,8)}

# # Extract residuals and fitted values
# residuals <- resid(model)
# fitted <- fitted(model)
# 
# # Plot residuals vs. fitted values
# ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = Residuals)) +
#   geom_point() +
#   geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
#   labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals")

```

## LME. Model quality checks

```{r, fig.dim = c(8,8)}
# ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = sqrt(abs(Residuals)))) +
#   geom_point() +
#   geom_smooth(method = "loess", color = "red") +
#   labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Square Root of |Residuals|")
```

## LME. Model quality checks

```{r}
# model_reduced <- lmerTest::lmer(ResponseNorm ~ Intervention + Time + (1|Subject/Intervention),
#                      data = FCdata, REML = T,
#                      control = lmerControl(optCtrl = list(maxfun = 1e6)))
# check_collinearity(model_reduced)
```

## Time*Intervention interaction plot: Data

```{r, fig.dim = c(10,8)}

sub_sum_FCdata <- FCdata %>%   
  group_by(Intervention, Time, Subject) %>%  
  summarize(FC_normed_mean = mean(ResponseNorm)) %>%
  ungroup()

sum_FCdata <- sub_sum_FCdata %>%   
  group_by(Intervention, Time) %>%  
  summarize(ResponseNorm_mean = mean(FC_normed_mean),
            ResponseNorm_se = sd(FC_normed_mean)/sqrt(length(FC_normed_mean))) %>%
  ungroup()

sum_FCdata %>%
  ggplot(aes(x = Time,  
             y = ResponseNorm_mean,
             group = Intervention,  
             color = Intervention)) +    
  geom_point(size = 3.5, position = position_dodge(width = 0.2)) + 
  geom_line(size=1.5, position = position_dodge(width = 0.2)) +
  ylab("Mean normalised FC") +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se,
                    ymin=ResponseNorm_mean-ResponseNorm_se), width=0, size = 1,
                position = position_dodge(width = 0.2)) +
  scale_colour_manual(values=c("#999999", "#663333", "#004488", "#DDAA33")) +
  scale_x_discrete(labels=c("Base 2", "Base 1", "0 min", "30 min", "60 min"))
```

## Time*Intervention interaction plot: Model prediction

```{r, echo=FALSE}
emm.model <- emmeans(model, ~ Intervention * Time,  data = FCdata)

emmip_output <- emmip(emm.model, ~ Intervention*Time ,type = "response", data = FCdata, CIs = TRUE, plotit=FALSE)

emmip_output %>%
  ggplot(aes(x=Time, 
             y=yvar,
             group = Intervention,  
             color = Intervention)) +
  geom_point(size = 3.5, position = position_dodge(width = 0.2)) + 
  geom_line(size=1.5, position = position_dodge(width = 0.2)) +
  ylab("Predicted response")+
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=0, size = 1,
                position = position_dodge(width = 0.2)) +
  scale_colour_manual(values=c("#999999", "#663333", "#004488", "#DDAA33")) +
  theme(legend.position="top") +
  scale_x_discrete(labels=c("Base 2", "Base 1", "0 min", "30 min", "60 min"))

```

## Pairwise comparisons: Time*Intervention interaction

```{r}
emm_options(lmer.df = 'asymptotic') 

emm.model <- emmeans(model, ~ Intervention * Time, data = FCdata)

emm.model.contrast = contrast(emm.model, "trt.vs.ctrl1", simple = "each",
                                   combine = TRUE, adjust = "fdr")

contrast.summary <- summary(emm.model.contrast,by = NULL)
contrast_pretty <- kable(get_anova_table(contrast.summary), digits = 3)
print(contrast_pretty)

write.csv(contrast.summary, file = "contrasts_FC.csv")
save_kable(contrast_pretty,"contrasts_FC.txt")
```
