---
title: "FCAnalysisAEC"
output: powerpoint_presentation
date: "2024-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, echo=FALSE, warning = FALSE, message=FALSE}
library(tidyverse)
library(lattice)
library(latticeExtra)
library(gridExtra) # adjust plots in a grid
library(lme4) # linear mixed models
library(kableExtra)
library(nlme)
library(performance)
library(emmeans) # post hoc multiple comparisons of linear mixed models
library(see)
library(patchwork)
```

```{r, echo=FALSE}
setwd("W:/Projects/2019-04 M1M1PAS Project/analysis/M1M1PASgit")

FCdata <- read.csv('W:/Projects/2019-04 M1M1PAS Project/analysis/source_Paolo/fc_source/sourceConnectivityBeamformeroldLF.csv', header=TRUE, sep=',', dec='.')
FCdata$Subject <- factor(FCdata$subject)
FCdata$Intervention <- factor(FCdata$task)
FCdata$Time <- factor(FCdata$run,levels=c("1","2","3","4","5"))
FCdata$wPLI <- as.numeric(FCdata$wpli)
FCdata$ciPLV <- as.numeric(FCdata$ciplv)
FCdata$AEC <- as.numeric(FCdata$aec)
```

```{r, echo=FALSE}

FCdata$Response <- FCdata$wPLI

```
## Check distribution of the dependant variable

```{r, echo=FALSE, fig.dim = c(12,8)}

xy.time <- xyplot(Response ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "FC", lwd = 2)

xy.interv <- xyplot(Response ~ Intervention, data = FCdata, type = c("a", "p"), pch = 20,
                 group = Subject, xlab = "Intervention", ylab = "FC", lwd = 2)
                    
qq.data <- qqmath(~ Response, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

grid.arrange(grobs = list(xy.time,xy.interv,qq.data), ncol = 2, nrow = 2)
```
## Check distribution of the dependant variable

```{r, echo=FALSE}
hist(FCdata$Response, xlab = "FC")
```

The AEC is not normally distributed.

## Check distribution of the normalised variable

We transform it by taking it to the power of 0.30.

```{r, echo=FALSE}
FCdata$ResponseNorm <- atanh(FCdata$Response)^0.24
```
## Check distribution of the normalised variable

```{r, echo=FALSE, fig.dim = c(12,8)}
xy.time <- xyplot(ResponseNorm ~ Time, data = FCdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "Normalised FC", lwd = 2)

xy.interv <- xyplot(ResponseNorm ~ Intervention, data = FCdata, type = c("a", "p"), pch = 
                      20, group = Subject, xlab = "Intervention", ylab = "Normalised FC", 
                    lwd = 2)
                    
qq.data <- qqmath(~ ResponseNorm, data = FCdata,
                  ylab = "FC",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

grid.arrange(grobs = list(xy.time,xy.interv,qq.data), ncol = 2, nrow = 2)

```
## Check distribution of the normalised variable

```{r, echo=FALSE}
#FCdata$ResponseNorm <- (FCdata$Response)^0.27
#FCdata$ResponseNorm <- atanh(FCdata$Response)^0.24
#FCdata$ResponseNorm <- atanh(FCdata$Response)^0.24
hist(FCdata$ResponseNorm, xlab = "normalised FC")

```
## LME. Model selection

```{r}
model1 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

model2 <- lme4::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

anova(model1,model2)
```
## LME. Model selection

A model with a session as a random effect (1 \| Subject/Intervention) has lower AIC and BIC values, so we will take it for further analysis.

## LME. Significance test

```{r}
model <- lmerTest::lmer(ResponseNorm ~ Intervention*Time + (1|Subject/Intervention),
                     data = FCdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
car::Anova(model)
```
## LME. Significance test

2-way interaction of Intervention:Time is significant. The main effect of Time is also significant.

## LME. Model quality checks

```{r, echo=FALSE, fig.dim = c(15,30)}
check_model(model)
```
## LME. Model quality checks

Model residuals are approximately okay but collinearity is very high.

## LME. Model quality checks

```{r, echo=FALSE}
library(ggplot2)

# Extract residuals and fitted values
residuals <- resid(model)
fitted <- fitted(model)

# Plot residuals vs. fitted values
ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = Residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs. Fitted Values", x = "Fitted Values", y = "Residuals")

```
## LME. Model quality checks

```{r, echo=FALSE}
ggplot(data.frame(Fitted = fitted, Residuals = residuals), aes(x = Fitted, y = sqrt(abs(Residuals)))) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue") +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Square Root of |Residuals|")
```
## LME. Model quality checks

```{r}
model_reduced <- lmerTest::lmer(ResponseNorm ~ Intervention + Time + (1|Subject/Intervention),
                     data = FCdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
check_collinearity(model_reduced)
```

If we fit a model without interactions, with just main effects, there is no collinearity. So we will ignore this issue.

## Post-hoc plots of significant effects

We will plot not raw but normalised data, the way it was used in the model. Since it's approximately normal, the y-axis is mean and SE of the data.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.dim = c(12,8)}
sub_sum_noInt_FCdata <- FCdata %>%   
  group_by(Subject, Time) %>%  
  summarize(ResponseNorm_subject_mean = mean(ResponseNorm)) %>%
  ungroup()

sum_noInt_FCdata <- sub_sum_noInt_FCdata %>%   
  group_by(Time) %>%  
  summarize(ResponseNorm_mean = mean(ResponseNorm_subject_mean),
            ResponseNorm_se = sd(ResponseNorm_subject_mean)/sqrt(length(ResponseNorm_subject_mean))) %>%
  ungroup()

sum_noInt_FCdata %>%
  ggplot(aes(x = as.numeric(Time),  
             y = ResponseNorm_mean)) +    
  geom_point(size = 3) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se, ymin=ResponseNorm_mean-ResponseNorm_se), width=.1, size = 1) 
  
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.dim = c(12,8)}

sum_FCdata <- FCdata %>%   
  group_by(Intervention, Time) %>%  
  summarize(ResponseNorm_mean = mean(ResponseNorm),
            ResponseNorm_se = sd(ResponseNorm)/sqrt(length(ResponseNorm))) %>%
  ungroup()

sum_FCdata %>%
  ggplot(aes(x = Time,  
             y = ResponseNorm_mean,
             group = Intervention,  
             color = Intervention)) +    
  geom_point(size = 3) + 
  geom_line(size=1) +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se, ymin=ResponseNorm_mean-ResponseNorm_se), width=.1, size = 1) +
  theme(legend.position="top")
```

## Post-hoc pairwise comparisons

```{r, echo=FALSE}

emm.model <- emmeans(model, ~ Time, data = FCdata)

emm.model.contrast = contrast(emm.model, "trt.vs.ctrl1", 
                                   combine = TRUE, adjust = "fdr")

contrast.summary <- summary(emm.model.contrast,by = NULL)
contrast.summary

```
## Model prediction
```{r, echo=FALSE}

emmip_output <- emmip(emm.model, ~ Time ,type = "response", data = FCdata, CIs = TRUE, plotit=FALSE)

ggplot(data = emmip_output, aes(x=Time, y=yvar)) +
  geom_point(size = 2) + 
  geom_line(group = 1, size = 0.8) +
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=.06, size = 0.8) +
  ylab("Predicted response")

```
## Post-hoc pairwise comparisons

```{r, echo=FALSE}
#emm_options(lmerTest.limit = 23040)
#emm_options(pbkrtest.limit = 23040)
emm_options(lmer.df = 'asymptotic') # also possible: 'satterthwaite', 'kenward-roger', 'asymptotic'

emm.model <- emmeans(model, ~ Intervention * Time, data = FCdata)

emm.model.contrast = contrast(emm.model, "pairwise", simple = "each",
                                   combine = TRUE, adjust = "fdr")

summary(emm.model.contrast,by = NULL)
```
## Model prediction

```{r, echo=FALSE}
emm.model <- emmeans(model, ~ Intervention * Time,  data = FCdata)

emmip_output <- emmip(emm.model, ~ Intervention*Time ,type = "response", data = FCdata, CIs = TRUE, plotit=FALSE)

ggplot(data = emmip_output, aes(x=Time, y=yvar,group = Intervention,  
             color = Intervention)) +
  geom_point(size = 2) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=.06, size = 0.8) +
  #facet_grid(~Intervention) +
  #scale_color_manual(values = c("grey47", "black", "black", "black", "black")) +
  ylab("Predicted response")

```

```{r}
