---
title: "SIHI analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r libraries}
library(tidyverse)
library(lattice) #plotting
library(gridExtra) #plotting
library(lme4) #LME
library(performance) # model evaluation
library(emmeans) #pairwise comparisons
library(kableExtra) #pretty tables
library(rstatix) #pretty tables
```

```{r}
setwd("W:/Projects/2019-04 M1M1PAS Project/analysis/M1M1PASgit")
MEPdata <- read.csv('MEPdata.csv', header=TRUE, sep=',', dec='.')
MEPdata$Subject <- factor(MEPdata$Subject)
MEPdata$Intervention <- factor(MEPdata$Intervention,levels=c("random","negneg","negpos","posneg"))
MEPdata$Time <- factor(MEPdata$Time,levels=c("Pre","0","30","60"))
MEPdata$Channel <- factor(MEPdata$Channel)
MEPdata$Intensity <- as.numeric(MEPdata$Intensity)
MEPdata$Response <- as.numeric(MEPdata$Response)
FDIdata <- subset(MEPdata, Channel == "FDIl")
```

## Check distribution of the dependant variable

```{r, fig.dim = c(15,10)}

xy.time <- xyplot(Response ~ Time, data = FDIdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "SIHI", lwd = 2)

xy.interv <- xyplot(Response ~ Intervention, data = FDIdata, type = c("a", "p"), pch = 20,
                 group = Subject, xlab = "Intervention", ylab = "SIHI", lwd = 2)

xy.intens <- xyplot(Response ~ Intensity, data = FDIdata, type = c("a", "p"), pch = 20,
                    group = Subject, xlab = "Intensity", ylab = "SIHI", lwd = 2)
                    
qq.data <- qqmath(~ Response, data = FDIdata,
                  ylab = "SIHI",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FDIdata, aes(x = Response)) +
  geom_histogram(binwidth = 1) +
  xlab("SIHI") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,xy.intens,qq.data,hist.data), ncol = 3, nrow = 2)
```

## Check distribution of the normalised variable

```{r, fig.dim = c(15,10)}
FDIdata$ResponseNorm <- (FDIdata$Response)^0.25

xy.time <- xyplot(ResponseNorm ~ Time, data = FDIdata, type = c("a", "p"), pch = 20,
                  group = Subject, xlab = "Time", ylab = "Normalised SIHI", lwd = 2)

xy.interv <- xyplot(ResponseNorm ~ Intervention, data = FDIdata, type = c("a", "p"), pch = 20,
                 group = Subject, xlab = "Intervention", ylab = "Normalised SIHI", lwd = 2)

xy.intens <- xyplot(ResponseNorm ~ Intensity, data = FDIdata, type = c("a", "p"), pch = 20,
                    group = Subject, xlab = "Intensity", ylab = "Normalised SIHI", lwd = 2)
                    
qq.data <- qqmath(~ ResponseNorm, data = FDIdata,
                  ylab = "Normalised SIHI",
                  prepanel = prepanel.qqmathline,
                  distribution = qnorm,
                  panel = function(x, ...) {
                    #panel.qqmathci(x, ...)
                    panel.qqmathline(x, ...)
                    panel.qqmath(x, ...)
                  })

hist.data <- ggplot(FDIdata, aes(x = ResponseNorm)) +
  geom_histogram(binwidth = 0.05) +
  xlab("Normalised SIHI") +
  theme_minimal()

grid.arrange(grobs = list(xy.time,xy.interv,xy.intens,qq.data,hist.data), ncol = 3, nrow = 2)

```

## LME. Model selection

```{r}
model1 <- lme4::lmer(ResponseNorm ~ Intervention*Time*Intensity + (1|Subject),
                     data = FDIdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

model2 <- lme4::lmer(ResponseNorm ~ Intervention*Time*Intensity + (1|Subject/Intervention),
                     data = FDIdata, REML = F,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))

anova(model1,model2)
```

## LME. Significance test

```{r}
model <- lmerTest::lmer(ResponseNorm ~ Intervention*Time*Intensity + (1|Subject/Intervention),
                     data = FDIdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
car::Anova(model)
```

## LME. Model quality checks

```{r, fig.dim = c(15,30)}
check_model(model)
```

## LME. Model quality checks

```{r}
model_reduced <- lmerTest::lmer(ResponseNorm ~ Intervention + Time + Intensity + (1|Subject/Intervention),
                     data = FDIdata, REML = T,
                     control = lmerControl(optCtrl = list(maxfun = 1e6)))
check_collinearity(model_reduced)
```

## Time*Intervention interaction plot: Data

```{r, fig.dim = c(10,8)}
sum_noInt_FDIdata <- FDIdata %>%   
  group_by(Intervention, Time, Subject) %>%  
  summarize(SIHI_normed_mean = mean(ResponseNorm)) %>%
  ungroup()

sub_sum_noInt_FDIdata <- sum_noInt_FDIdata %>%   
  group_by(Intervention, Time) %>%  
  summarize(ResponseNorm_mean = mean(SIHI_normed_mean),
            ResponseNorm_se = sd(SIHI_normed_mean)/sqrt(length(SIHI_normed_mean))) %>%
  ungroup()

sub_sum_noInt_FDIdata %>%
  ggplot(aes(x = Time,  
             y = ResponseNorm_mean,
             group = Intervention,  
             color = Intervention)) +    
  geom_point(size = 3.5, position = position_dodge(width = 0.2)) + 
  geom_line(size=1.5, position = position_dodge(width = 0.2)) +
  ylab("Mean normalised SIHI") +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se,
                    ymin=ResponseNorm_mean-ResponseNorm_se), width=0, size = 1,
                position = position_dodge(width = 0.2)) +
  scale_colour_manual(values=c("#999999", "#663333", "#004488", "#DDAA33")) +
  scale_x_discrete(labels=c("Baseline","0 min", "30 min", "60 min"))
  
```

## Time*Intervention interaction plot: Model prediction

```{r, fig.dim = c(10,8)}
emm.model <- emmeans(model, ~ Intervention * Time,  data = FDIdata)

emmip_output <- emmip(emm.model, ~ Intervention*Time ,type = "response", data = FDIdata, CIs = TRUE, plotit=FALSE)

emmip_output %>%
  ggplot(aes(x=Time, 
             y=yvar,
             group = Intervention,  
             color = Intervention)) +
  geom_point(size = 3.5, position = position_dodge(width = 0.2)) + 
  geom_line(size=1.5, position = position_dodge(width = 0.2)) +
  ylab("Predicted response")+
  geom_errorbar(aes(ymax=UCL, ymin=LCL), width=0, size = 1,
                position = position_dodge(width = 0.2)) +
  scale_colour_manual(values=c("#999999", "#663333", "#004488", "#DDAA33")) +
  theme(legend.position="top") +
  scale_x_discrete(labels=c("Baseline","0 min", "30 min", "60 min"))

```

## Pairwise comparisons: Time*Intervention interaction

```{r}
emm_options(lmer.df = 'asymptotic')

emm.model <- emmeans(model, ~ Intervention * Time, data = FDIdata)

emm.model.contrast = contrast(emm.model, "trt.vs.ctrl1", simple = "each",
                                   combine = TRUE, adjust = "fdr")

contrast.summary <- summary(emm.model.contrast,by = NULL)
contrast_pretty <- kable(get_anova_table(contrast.summary), digits = 3)
print(contrast_pretty)

write.csv(contrast.summary, file = "contrasts_SIHI.csv")
save_kable(contrast_pretty,"contrasts_SIHI.txt")
```

## Intervention*Intensity interaction: Data

```{r, fig.dim = c(10,8)}
sum_noTime_FDIdata <- FDIdata %>%   
  group_by(Intervention, Intensity, Subject) %>%  
  summarize(SIHI_normed_mean = mean(ResponseNorm)) %>%
  ungroup()

sub_sum_noTime_FDIdata <- sum_noTime_FDIdata %>%   
  group_by(Intervention, Intensity) %>%  
  summarize(ResponseNorm_mean = mean(SIHI_normed_mean),
            ResponseNorm_se = sd(SIHI_normed_mean)/sqrt(length(SIHI_normed_mean))) %>%
  ungroup()

sub_sum_noTime_FDIdata %>%
  ggplot(aes(x = Intensity,  
             y = ResponseNorm_mean,
             group = Intervention,  
             color = Intervention)) +    
  geom_point(size = 3.5, position = position_dodge(width = 0.2)) + 
  geom_line(size=1.5, position = position_dodge(width = 0.2)) +
  ylab("Mean normalised SIHI") +
  scale_colour_manual(values=c("#999999", "#663333", "#004488", "#DDAA33")) +
  geom_errorbar(aes(ymax=ResponseNorm_mean+ResponseNorm_se, ymin=ResponseNorm_mean-ResponseNorm_se), width=.1, size = 1, position = position_dodge(width = 0.2)) +
  theme(legend.position="top")
```
